Real Time Variational Density Forecasts

### 2017-01-30 Confirmation/confirmation_V2.tex
confirmation 2nd draft

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{wasysym}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\numberwithin{equation}{section}
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\tableofcontents
\section{Introduction}

@@

As a contrast to frequentism, the Bayesian methodology has been used extensively in forecasting \citep{Geweke2006} and is of interest as the entire probability density for $p(y_{T+1} | y_{1:T})$ is provided implicitly. However the Bayesian approach requires solving integrals of the form

\begin{equation}
\label{forecastintegral}
\int_\theta p(y_{T+1} | \theta) p(\theta | y_{1:T}) d \theta,
\end{equation}

where $p(\theta | y_{1:T})$ is known as the posterior distribution, with

\begin{equation}
\label{posterior}
 p(\theta | y_{1:T}) = \frac{p(y_{1:T}|\theta)p(\theta)}{\int_\theta p(y_{1:T}|\theta)p(\theta) d\theta}
\end{equation}

and $p(\theta)$ is some pre-specified prior distribution.

These often cannot be solved analytically, while numeric integration is computationally infeasible when the dimension of $\theta$ is large. To address this problem there is a wide range of techniques used to approximate the solution to the integral in (\ref{forecastintegral}), such as Markov Chain Monte Carlo (MCMC) and Variational Bayes (VB). These approximations have an implicit trade-off: The better approximations are compuationaly intensive, and the forecaster must decide how much computation time and approximation error is acceptable. The focus of this research is in situtations where MCMC is feasible in a moderate amount of time, but is infeasible in time period between two observations. To move $p(y_{T+1} | y_{1:T})$ to $p(y_{T+2} | y_{1:T+1})$ requires updating the posterior distribution after observation of each new data point observed via

\begin{equation}
\label{posteriorupdate}
p(\theta | y_{1:T+1}) = \frac{p(y_{T+1} | \theta) p(\theta | y_{1:T})}{\int_\theta p(y_{T+1} | \theta) p(\theta | y_{1:T}) d\theta}.
\end{equation}

However, if the data is based on a market such as with stock returns the market will close overnight and MCMC can be ran.

\section{Bayesian Inference}
\subsection{Exact Bayesian Computation}
While it is technically an approximation method, MCMC algorithms result in what is often called an exact computation of the posterior. A Gibbs based MCMC iteratively samples from the conditional distributions

\begin{align}
&p(\theta_1 | \theta_2, \dots, \theta_p, y_{1:T}) \nonumber \\
&p(\theta_2 | \theta_1, \theta_3, \dots, \theta_p, y_{1:T}) \nonumber \\
&\vdots \nonumber \\
&p(\theta_p | \theta_1, \dots, \theta_{p-1}, y_{1:T}) \nonumber
\end{align}

where $p$ is the dimension of $\theta$. With enough iterations, the error in the approximation converges to zero and the algorithm can be ran for as much time as the forecaster desires to reduce error to a desired level. However, in the first iteration we must set arbitary starting values for each of $\theta_2, \dots, \theta_p$ introducing a large amount of error in the early iterations. To avoid this error MCMC generally must be run for a large number of iterations and these early samples are discarded. The computation time per iteration and speed of convergence is extremely problem specific.

We illustrate this with an AR(2), a simple time series model used in the remainder of this section. The AR(2) is described by

\begin{equation}
\label{AR2}
y_t - \mu = \rho_1 (y_{t-1} - \mu) + \rho_2 (y_{t-2} - \mu) + \epsilon_t
\end{equation}

where $\epsilon_t \sim \mathcal{N}(0, \sigma^2)$. We collect the unknown parameters as $\theta = (\mu, \sigma^2, \rho_1, \rho_2)'$ and set priors as

>>>>

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\documentclass[12pt,a4paper]{article}%
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{amsmath}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{url}
\usepackage{wasysym}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\setcounter{MaxMatrixCols}{30}
\providecommand{\U}[1]{\protect\rule{.1in}{.1in}}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\  \rule{0.5em}{0.5em}}
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\evensidemargin}{0.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.25in}
\title{Real Time Variational Density Forecasts}
\author{Nathaniel Tomasetti, Catherine Forbes and Anastasios Panagiotelis}
\begin{document}

\maketitle
\tableofcontents
\section{Introduction}

@@

As a contrast to frequentism, the Bayesian methodology has been used extensively in forecasting \citep{Geweke2006} and is it can capture the error components from parameter estimation and modelling error. However calculation of the Bayesian posterior distribution, $p(\theta | y)$ requires solving complex integrals that often do not have an analytical solution, in addition, if the the dimension of the parameter vector $\theta$ is large then numerical integration will be computationally infeasible.

To address this problem a wide range of techniques are available, such as Markov Chain Monte Carlo (MCMC) and Variational Bayes (VB). The forecasters decision of technique to use is a compromise to implicit trade-off: MCMC captures the posterior distribution accurately but is compuationaly intensive, while other methods such as VB tend to be faster but are inaccurate. One time series of interest, electricity demand data, operates on five minute interval in Victoria, too short for MCMC to be a feasible option in a moderately complex model. The focus of this research is in time series that operate in real-time, where the observation of each data point requires the model parameters to be updated before a density forecast of the next data point can be made. Variational Bayes is used as an alternative computation strategy, and the trade off between the statistical and computational efficiency will be explored using data on the electricity demand of Victoria for the 2012-2015 period.

\section{Bayesian Inference}

Exact Bayesian inference in the context of forecasting uses the predictive density of $Y_{T+1}$ conditioned on the observed series, $Y_{t}, t = 1, \dots T$, which is denoted by $y_{1:T}$. The first step requires the posterior distribution of the model parameters $\theta$, p(\theta | y), where

\begin{equation}
\label{posterior}
 p(\theta | y_{1:T}) = \frac{p(y_{1:T}|\theta)p(\theta)}{\int_\theta p(y_{1:T}|\theta)p(\theta) d\theta}.
\end{equation}

Generally the solution to (\ref{posterior}) is unavailable. This section will offer two alternatives, MCMC and VB.
MCMC is used to create a sample from $p(\theta | y)$, and any function of interest from $p(\theta | y$ can be found with the relevant function applied to the sample, while VB replaces $p(\theta | y)$ with an approximation $q(\theta | \lambda)$, choosing some family $q$ and approximation parameters $\lambda$.

\subsection{Markov Chain Monte Carlo}

There are many types of MCMC algorithm, the most common being a Gibbs sampler, which iteratively samples from a $p$ dimensional $\theta$ via the conditional distributions

\begin{align}
&p(\theta_1 | \theta_2, \dots, \theta_p, y_{1:T}) \nonumber \\
&p(\theta_2 | \theta_1, \theta_3, \dots, \theta_p, y_{1:T}) \nonumber \\
&\vdots \nonumber \\
&p(\theta_p | \theta_1, \dots, \theta_{p-1}, y_{1:T}). \nonumber
\end{align}

With enough iterations, these samples converge in distribution to $p(\theta | y)$ and can be used for inference. The computation time per iteration and number of iterations required for convergence is problem specific but typically increases with model complexity.

To illustrate issues involved consider a simple auto-regressive time series model which will be used in the remainder of this section. This model is described by

\begin{equation}
\label{AR2}
y_t = \rho_1 y_{t-1} + \rho_2 y_{t-2} + \epsilon_t
\end{equation}

where $\epsilon_t \sim \mathcal{N}(0, \sigma^2)$. The likelihood of parameters $\theta = (\rho_1, \rho_2m \sigma^2)'$ is given by

\begin{align}
\label{likelihood}
L(\theta | y_{1:T}) &= p(y_1 | \theta) p(y_2 | y_1, \theta) \prod_{t=3}^{T}p(y_t | y_{1:t-1}, \theta) \nonumber \\
&= \frac{1}{(2\pi)^{T/2}\gamma(0)^{1/2}\Gamma{0}^{1/2}\sigma^{(T-2)/2}} \exp \left\{ \frac{-y_1^2}{2 \gamma(0)} \right\} 
\exp \left\{\frac{-(y_2 - \gamma(1)/\gamma(2)y_1)^2}{2 \Gamma(0)} \right} \exp \left\{ \frac{-\sum_{t=3}^{T}(y_t - \phi_1 y_{t-1} - \phi_2 y_{t-2})^2}{2\sigma^2} \right\}
\end{align}

where $\gamma(k)$ is the $k-th$ autocorrelation and $\Gamma(1) = \gamma(0) - \gamma(1)^2 / \gamma(0)$.
